---
title: "Analysis Pipeline"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Loading Libraries
```{r}
library(usdm)
library(dplyr)
library(ape)
library(phylolm)
library(caret)
library(corrplot)
```

#1 Importing Data

##1.1 Domesticate Niche trait Data
###Load in the Plant (angiosperm) and Ungulate niche trait occurence data.
###Rename Species column name and name every row after the species name.
```{r}
#loading plant trait data
plant.data.location <- "https://raw.githubusercontent.com/rvosa/sdmdl-angiosperm-data/master/data/niche_traits_merged.csv"
plant.data.df <- read.csv(plant.data.location)
names(plant.data.df)[names(plant.data.df) == "allmb_name"] <- "Species"

#renaming rows after speices name
row.names(plant.data.df) <- plant.data.df$Species

#loading ungulate trait data
ungulate.data.location <- "https://raw.githubusercontent.com/naturalis/trait-geo-diverse-ungulates/master/results/OMI/niche_traits.csv"
ungulate.data.df <- read.csv(ungulate.data.location)
names(ungulate.data.df)[names(ungulate.data.df) == "X"] <- "Species"

#renaming rows after species name
row.names(ungulate.data.df) <- ungulate.data.df$Species
```

###Joining angiosperm and ungulate niche trait data to give a combined dataframe of domesticate niche traits. 
```{r}
combined.niche.trait.data <- rbind(plant.data.df, ungulate.data.df)
```


##1.2 Societal Domesticate-Portfolio data
###Load normalised species presence data for each society. 
###This table is based on variables taken from D-PLACE (see "/subsistence-comparison/data/Societies/normalised_sccs_vars_for_states.csv").
###Definitions of the states seen in this table can be found at ("/subsistence-comparison/data/Societies/normalised_state_definitions.csv").
```{r}
domesticate.portfolio.df <- read.csv("../data/Societies/normalised_species_presence.csv")
```


##1.3 Join combined.niche.trait.data with domesticate.portfolio.df, on column "species"
###This gives the niche trait information for all environmental variables and the societies within which they occure for all plant and ungulate domesticated species.
###The domesticate.niche.data.df is used externally to calculate societal niche traits for every environmental variable from a weighted average of the niche traits of their domesticates (see below). 
```{r}
domesticate.niche.data.df <- merge(domesticate.portfolio.df, combined.niche.trait.data, sort = F,)

row.names(domesticate.niche.data.df) <- domesticate.niche.data.df$Species

write.csv(domesticate.niche.data.df, file = "../data/domesticate_niche_data_df.csv")
```

##1.4 Load Societal niche traits
###Read the Instructions for calculating the envirnomental traits for each society, based on the weighted niche traits of their domesticate portfolio, can be found in the documents "README_calculating_societal_niche_traits.docx", and the accompanying "domesticate_niche_data.xlsx". The result of this calculation can be found in "societal_niche_traits.csv" (see below).

###This calculation also uses the A:H ratio data (The societal dependence on agriculture/animal husbandry table is used for weightings calculations (to follow). The ratios have been derived from D-PLACE variables: SCCS203-SCCS207.). See: "../data/Societies/relative_weightings_H-A.csv". 

###Load the "societal_niche_traits.csv" for further analysis.
###The column containing the society SCCS codes is also renamed "Society".
```{r}
societal.niche.traits.location <- "https://raw.githubusercontent.com/naturalis/subsistence-comparison/master/data/societal_niche_traits.csv"

#making a data frame of the niche traits for each society
societal.niche.traits.df <- read.csv(societal.niche.traits.location)

#renaming the X column "Society"
names(societal.niche.traits.df)[names(societal.niche.traits.df) == "X"] <- "Society"
```


##1.5 Load basic information on the SCCS societies (soc.info.df)
###Basic information will include D_PLACE society number, society name and Glottocode
```{r}
all.soc.info.df <- read.csv("../data/Societies/Basic_Society_Information.csv")
#removing societies that do not primarily depend upon animal Husbandry or Agriculture for subsistence (see "relative_weightings_H-A.csv").
relevant.soc.info.df <- (all.soc.info.df[-c(1,3,5,7,10:12,22:23,25:45,47,49:55,58:59,64,70:71,73:76,80,83:89,91:94,96:97,99,136,162,164,166,172,176:178,182:184),])
#removing columns of extra information
soc.info.df <- (relevant.soc.info.df[,-c(2,4,6:11)])
#renaming rows after D-PLACE code
row.names(soc.info.df) <- soc.info.df$D.PLACE_id
```


##1.6 Load in language tree.
###Import original phylogeny
###Prune tree
```{r}
#Import original phylogeny (tree) 
##The global glottolog tree is imported fro the D-PLACE git repository
tree <- read.nexus("https://raw.githubusercontent.com/D-PLACE/dplace-data/master/phylogenies/glottolog_global/summary.trees")

#remove those branches of the tree that do not correspond with the list of glottocodes for the 113 societies. Call this new relevant tree Language.tree
tips.to.drop <- setdiff(tree$tip.label, soc.info.df$Glottocode)
language.tree <- drop.tip(tree, tips.to.drop)

#view the language tree, to see if it looks reasonable. 
plot(language.tree)
```


```{r}
#Viewing the labels vector below, we see that there are only 106 societies listed, not the 113 we were hoping for. It seems 7 glottocodes are missing from the tree.
labels <- language.tree$tip.label

#The difference between the total list of 113 glottocodes from our society list and those that have matches with the language tree, gives us the 7 missing glottocodes from the tree (see: missing), these must be removed from the analysis.
missing <- setdiff(soc.info.df$Glottocode, labels)
missing
```


#2 Processessing and Cleaning the data
###The society data used for the analysis must be edited to removed those societies that do not feature in tree.
###Leaving a total of 106 societies for further analysis
```{r}
#rename D-PLACE_id to Society
names(soc.info.df)[names(soc.info.df) == "D.PLACE_id"] <- "Society"
#join soc.info.df to societal.niche.traits.df on column Society
societies.df <- cbind(soc.info.df, societal.niche.traits.df[,c(2:42)])

#remove the rows that are not included in the tree (i.e. those listed under "missing")
societies.df <- societies.df[!grepl("kore1280", societies.df$Glottocode),]
societies.df <- societies.df[!grepl("nucl1417", societies.df$Glottocode),]
societies.df <- societies.df[!grepl("cusc1236", societies.df$Glottocode),]
societies.df <- societies.df[!grepl("mapu1245", societies.df$Glottocode),]
societies.df <- societies.df[!grepl("tobe1252", societies.df$Glottocode),]
societies.df <- societies.df[!grepl("abui1241", societies.df$Glottocode),]
societies.df <- societies.df[!grepl("siwa1245", societies.df$Glottocode),]
#check that all remaining societies are in the language tree. Check should be empty.
check <- setdiff(societies.df$Glottocode, labels)
check
```

```{r}
#cleaning up a little
rm(check,all.soc.info.df,relevant.soc.info.df,soc.info.df,societal.niche.traits.df,missing,labels,tree)
```



#3 VIF analysis
##3.1 Estimating best (least collinear) variables using pairwise comparisons, before comparing these to a VIF
```{r}
###create initialPredictors containing the societies.df
initialPredictors <- societies.df[,c(4:44)]
###load this data into a matrix
matrix.all <- cor(initialPredictors, use = "pairwise.complete.obs")
###show the full corrplot pairwise comparison matrix for all 41 environmental variables (on the above matrix) 
corrplot(matrix.all, type="lower", order = "hclust", tl.pos = "l", tl.col = "black", tl.cex = 0.5)

#Using findCorrelation() to identify the environmental variables with the highest collinearity in the above matrix. 
cutoffValues <- c(0.9, 0.8, 0.7)
for(cut in cutoffValues){
  matrix <- cor(initialPredictors, use = "pairwise.complete.obs")
  delColumn <- findCorrelation(matrix, cutoff = cut, verbose = FALSE, exact=FALSE)
  initialPredictors <- initialPredictors[,-c(delColumn)]
}
vif(initialPredictors)

##Repeating the steps above with cut of values of .9 down to .6, it can be found that a cutoff between 0.7 and 0.8 is optimal. Using 0.8 gives 14 variables, of which 9 present a VIF over 10, while using a cutoff of .7 yields only 8 environmental variables, but all are under 4.
###The most promising environmental variables are: Aspect, BulkDensity, ClayPercentage, PETWettestQuarter, PhCaCL, Slope, bio14, bio19. With the variables: annualPET, minTempWarmest, PETDriestQuarter, bio2, bio15, bio18, being of potential but slightly lesser promise (appearing under a cut off of 0.8).
```

##3.2 Calculating collinearity using the Varibale Inflation Factor (VIF) for Environmental variable selection.
```{r}
#run VIFstep on the societal niche data
vifstep(societies.df[,c(4:44)],th=10)
```


###The resulting variables from the VIFstep are very similar to those shown to be the most promising (least collinear) variables form the pairwise analysis. (See a comparison of their matrices below.)
```{r}
#creating data frame of societies with niche trait columns of only those 12 environmental variables with VIF scores below 10, from the above VIF
VIF.variables.df <- (societies.df[,c(1:6,18:19,23:25,27,40,43:44)])
#making correlation matrix of all Environmental Variables for visual comparison
cor.tab <- cor(VIF.variables.df[,-c(1:3)])
```

##3.3 Comparing VIFstep and pairwise comparison analyses
###Using corrplot to visually compare the collinearity of the environmental variables between the results of the pairwise analysis "matrix", and the "cor.tab"  produced by the VIFstep
```{r}
#for pairwise
corrplot(matrix, type="lower", order = "hclust", tl.pos = "l", tl.col = "black", tl.cex = 0.5)
```


```{r}
#for VIF
corrplot(cor.tab, type="lower", order = "hclust", tl.pos = "l", tl.col = "black", tl.cex = 0.5)
```



#4 Model Selection/Analysis
##4.1 Selecting/prepping D-PLACE variables for comparison
```{r}
#Run the file: to load in all the D-PLACE variables and sort the data ready for model selection:
source("D-PLACE.variable.cleaning.R")
```


##4.2 Model Selection
```{r}
#formula for phyloglmstep/phylostep (the same for all D-PLACE variables being tested, as code represents the values of each dependent (societal) variable within that variable's data frame)
formula <- code ~ Aspect + BulkDensity + ClayPercentage + PETDriestQuarter + PETseasonality + OrganicCarbon + PhCaCL + Slope + bio2 + bio15 + bio18 + bio19
```


###4.2.i writing (binary)
```{r}
#writing.data phyloglmstep
phyloglmstep (formula, starting.formula = NULL, data=writing.data, phy = language.tree, method= "logistic_MPLE", direction = c("both"), trace = 2, btol = 100, log.alpha.bound = 4, start.beta=NULL, start.alpha=NULL, boot = 0, full.matrix = TRUE, k=2)

#The best model produced by the above phyloglmstep for writing.data:
writing.formula <- code ~ 1 + Aspect + BulkDensity + PETseasonality + OrganicCarbon + bio15
#AIC: 109.35 
#Alpha: 0.387

#running the GLM for writing.data
GLM.writing <- phyloglm(formula = writing.formula, data = writing.data, phy = language.tree, method = "logistic_MPLE", btol = 100, log.alpha.bound = 4)

summary(GLM.writing)
```


###4.2.ii money (binary)
```{r}
#money.data phyloglmstep
phyloglmstep (formula, starting.formula = NULL, data=money.data, phy = language.tree, method= "logistic_MPLE", direction = c("both"), trace = 2, btol = 100, log.alpha.bound = 4, start.beta=NULL, start.alpha=NULL, boot = 0, full.matrix = TRUE, k=2)

#The best model produced by the above phyloglmstep for money.data:
money.formula <- code ~ 1 + Aspect + PhCaCL + Slope
#AIC: 113.64
#Alpha: 2.470

#Running the GLM for money.data
GLM.money <- phyloglm(formula = money.formula, data = money.data, phy = language.tree, method = "logistic_MPLE", btol = 100, log.alpha.bound = 4)

summary(GLM.money)
```


###4.2.iii foreign.diseases (binary)
```{r}
#foreign.diseases.data phyloglmstep
phyloglmstep (formula, starting.formula = NULL, data=foreign.diseases.data, phy = language.tree, method= "logistic_MPLE", direction = c("both"), trace = 2, btol = 100, log.alpha.bound = 4, start.beta=NULL, start.alpha=NULL, boot = 0, full.matrix = T, k=2)

#The best model produced by the above phyloglmstep for foreign.diseases.data:
foreign.diseases.formula <- code ~ 1 + Aspect + ClayPercentage
#AIC: 84.63
#Alpha: 4.074

#Running the GLM for foreign.diseases
GLM.foreign.diseases <- phyloglm(formula = foreign.diseases.formula, data = foreign.diseases.data, phy = language.tree, method = "logistic_MPLE", btol = 100, log.alpha.bound = 4)

summary(GLM.foreign.diseases)
```


###4.2.iv warfare (continuous)
```{r}
#warfare.data phylostep
phylostep (formula, starting.formula = NULL, data=warfare.data, phy = language.tree, method= "logistic_MPLE", direction = "both", trace = 0, btol = 100, log.alpha.bound = 4, start.beta=NULL, start.alpha=NULL, boot = 0, full.matrix = TRUE, k=2)

#The best model produced by the above phylostep for warfare.data:
warfare.formula <- code ~ 1 + BulkDensity + OrganicCarbon + Slope + bio2 + bio15 + bio19
#AIC: 367.3
#Alpha: NA

#Running the LM for warfare.data
LM.warfare <- phylolm(formula = warfare.formula, data = warfare.data, phy = language.tree, method = "logistic_MPLE", btol = 100, log.alpha.bound = 4)

summary(LM.warfare)
```


###4.2.v community.size (continuous)
```{r}
#community.size.data phylostep
phylostep (formula, starting.formula = NULL, data=community.size.data, phy = language.tree, method= "logistic_MPLE", direction = c("both"), trace = 2, btol = 100, log.alpha.bound = 4, start.beta=NULL, start.alpha=NULL, boot = 10, full.matrix = TRUE, k=2)

#The best model produced by the above phylostep for community.size.data:
community.size.formula <- code ~ 1 + Aspect + BulkDensity + PETseasonality + OrganicCarbon + PhCaCL + bio15
#AIC: 1928.2
#Alpha: NA

#Running the LM for community.size.data
LM.community.size <- phylolm(formula = community.size.formula, data = community.size.data, phy = language.tree, method = "logistic_MPLE", btol = 100, log.alpha.bound = 4)

summary(LM.community.size)
```


###4.2.vi pop.density (continuous)
```{r}
#pop.density.data phylostep
phylostep (formula, starting.formula = NULL, data=pop.density.data, phy = language.tree, method= "logistic_MPLE", direction = c("both"), trace = 2, btol = 100, log.alpha.bound = 4, start.beta=NULL, start.alpha=NULL, boot = 0, full.matrix = TRUE, k=2)

#The best model produced by the above phylostep for pop.density.data:
pop.density.formula <- code ~ 1 + bio2 + bio15
#AIC: 1382.6
#Alpha: NA

#Running the LM for pop.density.data
LM.pop.density <- phylolm(formula = pop.density.formula, data = pop.density.data, phy = language.tree, method = "logistic_MPLE", btol = 100, log.alpha.bound = 4)

summary(LM.pop.density)
```


###4.2.vii pathogen.stress (continuous)
```{r}
#pathogen.stress.data phylostep
phylostep (formula, starting.formula = NULL, data=pathogen.stress.data, phy = language.tree, method= "logistic_MPLE", direction = c("both"), trace = 2, btol = 100, log.alpha.bound = 4, start.beta=NULL, start.alpha=NULL, boot = 0, full.matrix = TRUE, k=2)

#The best model produced by the above phylostep for pathogen.stress.data:
pathogen.stress.formula <- code ~ 1 + PETDriestQuarter + PETseasonality + PhCaCL + Slope + bio2
#AIC: 537.4
#Alpha: NA

#Running the LM for pathogen.stress.data
LM.pathogen.stress <- phylolm(formula = pathogen.stress.formula, data = pathogen.stress.data, phy = language.tree, method = "logistic_MPLE", btol = 100, log.alpha.bound = 4)

summary(LM.pathogen.stress)
```


###4.2.viii famine (binary)
```{r}
#famine.data phyloglmstep 
phyloglmstep (formula, starting.formula = NULL, data=famine.data, phy = language.tree, method= "logistic_MPLE", direction = c("both"), trace = 2, btol = 100, log.alpha.bound = 4, start.beta=NULL, start.alpha=NULL, boot = 0, full.matrix = T, k=2)

#The best model produced by the above phyloglmstep for famine.data:
famine.formula <- code ~ 1 + bio18
#AIC: 104.17
#Alpha: 2.719

#Running the GLM for famine
GLM.famine <- phyloglm(formula = famine.formula, data = famine.data, phy = language.tree, method = "logistic_MPLE", btol = 100, log.alpha.bound = 4)

summary(GLM.famine)
```


###4.2.ix political.integration (binary)
```{r}
#political.integration.data phyloglmstep 
phyloglmstep (formula, starting.formula = NULL, data=political.integration.data, phy = language.tree, method= "logistic_MPLE", direction = c("both"), trace = 2, btol = 100, log.alpha.bound = 4, start.beta=NULL, start.alpha=NULL, boot = 0, full.matrix = T, k=2)

#The best model produced by the above phyloglmstep for political.integration.data:
political.integration.formula <- code ~ 1 + ClayPercentage + PETseasonality + PhCaCL + Slope + bio18
#AIC: 115.48
#Alpha: 2.011

#Running the GLM for political.integration
GLM.political.integration <- phyloglm(formula = political.integration.formula, data = political.integration.data, phy = language.tree, method = "logistic_MPLE", btol = 100, log.alpha.bound = 4)

summary(GLM.political.integration)
```


###4.2.x social.strat (binary)
```{r}
#social.strat.data phyloglmstep
phyloglmstep (formula, starting.formula = NULL, data=social.strat.data, phy = language.tree, method= "logistic_MPLE", direction = c("both"), trace = 2, btol = 100, log.alpha.bound = 4, start.beta=NULL, start.alpha=NULL, boot = 0, full.matrix = T, k=2)

#The best model produced by the above phyloglmstep for social.strat.data:
social.strat.formula <- code ~ 1 + Aspect + PETseasonality + PhCaCL + bio15 + bio19
#AIC: 146.95
#Alpha: 1.052

#Running the GLM for social.strat
GLM.social.strat <- phyloglm(formula = social.strat.formula, data = social.strat.data, phy = language.tree, method = "logistic_MPLE", btol = 100, log.alpha.bound = 4)

summary(GLM.social.strat)
```


###4.2.xi violence (binary)
```{r}
#violence.data phyloglmstep
phyloglmstep (formula, starting.formula = NULL, data=violence.data, phy = language.tree, method= "logistic_MPLE", direction = c("both"), trace = 2, btol = 100, log.alpha.bound = 4, start.beta=NULL, start.alpha=NULL, boot = 0, full.matrix = T, k=2)

#The best model produced by the above phyloglmstep for violence.data:
violence.formula <- code ~ 1 + Aspect + PETDriestQuarter + PETseasonality + OrganicCarbon + PhCaCL + bio15 + bio19
#AIC: 104.05
#Alpha: 0.098

#Running the GLM for violence
GLM.violence <- phyloglm(formula = violence.formula, data = violence.data, phy = language.tree, method = "logistic_MPLE", btol = 100, log.alpha.bound = 4)

summary(GLM.violence)
```


###4.2.xii revenge (binary)
```{r}
#revenge.data phyloglmstep
phyloglmstep (formula, starting.formula = NULL, data=revenge.data, phy = language.tree, method= "logistic_MPLE", direction = c("both"), trace = 2, btol = 100, log.alpha.bound = 4, start.beta=NULL, start.alpha=NULL, boot = 0, full.matrix = T, k=2)

#The best model produced by the above phyloglmstep for revenge.data:
revenge.formula <- code ~ 1 + Aspect + BulkDensity + PETseasonality + PhCaCL + bio15
#AIC: 129.58
#Alpha: 4.821

#Running the GLM for revenge
GLM.revenge <- phyloglm(formula = revenge.formula, data = revenge.data, phy = language.tree, method = "logistic_MPLE", btol = 100, log.alpha.bound = 4)

summary(GLM.revenge)
```


###4.2.xiii defence (binary)
```{r}
#defence.data phyloglmstep
phyloglmstep (formula, starting.formula = NULL, data=defence.data, phy = language.tree, method= "logistic_MPLE", direction = c("both"), trace = 2, btol = 100, log.alpha.bound = 4, start.beta=NULL, start.alpha=NULL, boot = 0, full.matrix = T, k=2)

#The best model produced by the above phyloglmstep for defence.data:
defence.formula <- code ~ 1 + BulkDensity + PhCaCL + bio15 + bio18 + bio19
#AIC: 121.35
#Alpha: 0.664

#Running the GLM for defence
GLM.defence <- phyloglm(formula = defence.formula, data = defence.data, phy = language.tree, method = "logistic_MPLE", btol = 100, log.alpha.bound = 4)

summary(GLM.defence)
```


###4.2.xiv disasters (binary)
```{r}
#disasters.data phyloglmstep
phyloglmstep (formula, starting.formula = NULL, data=disasters.data, phy = language.tree, method= "logistic_MPLE", direction = c("both"), trace = 2, btol = 100, log.alpha.bound = 4, start.beta=NULL, start.alpha=NULL, boot = 0, full.matrix = T, k=2)

#The best model produced by the above phyloglmstep for disasters.data:
disasters.formula <- code ~ 1 + Aspect + ClayPercentage + PETDriestQuarter + Slope
#AIC: 62.62
#Alpha: 3.464

#Running the GLM for disasters
GLM.disasters <- phyloglm(formula = disasters.formula, data = disasters.data, phy = language.tree, method = "logistic_MPLE", btol = 100, log.alpha.bound = 4)

summary(GLM.disasters)
```



#5 Model Evaluation
##Testing the final models to ensure they still comply with GLM/LM prior-assumption requirements.

##This will be done using the package: "performance". Using the "check_model()" function. 
###(Note: The "check_model()" function is meant for models without a phylogeny, so we will be using GLM/LM instead of phyloglm/phylolm. However, also note the absence of the phylogeny here may skew the interpretation.
```{r}
library(performance)
library(see)
#library(gridExtra)
#library(car)
#library(randomForest)
#library(ResourceSelection)
```

```{r}
##writing model
writing.model.check <- glm(code ~ 1 + Aspect + BulkDensity + PETseasonality + OrganicCarbon + bio15, data = writing.data, family = binomial())

#model check overview
check_model(writing.model.check)
#No problems with multicollinearity
#No highly influential variables (Cook's Distance)
check_outliers(writing.model.check) #No outliers

#give summary of the model output
summary(writing.model.check)
#Null deviance: 142.35 on 105 df
#Residual deviance: 100.26 on 100 df

check_model(writing.model.check)
```


```{r}
##money
money.model.check <- glm(code ~ 1 + Aspect + PhCaCL + Slope, data = money.data, family = binomial())

#model checks
check_model(money.model.check)
#No multicollinearity problems
#No problematically high influential variables (Cook's Distance)
check_outliers(money.model.check) #no outliers

summary(money.model.check) #checking the deviance (null and residual)...

#Null deviance is 120.3 on 105 degrees of freedom, showing that the intercept (i.e. grand mean) does a pretty good job of predicting the response variable, this makes sense as the intercept here is highly significant.
#Residual deviance is 103.55 on 102 degrees of freedom, so the other variables improve the model's goodness of fit, at the cost of just 3 degrees of freedom.
#The model explains the ...

### The expected value of the chi^2 distribution is equal to its degree of freedom. A well fitting model will therefore show residual deviance that is approximately equal to its degrees of freedom. In other words, the "mean deviance" (i.e. deviance/df) should be approximately 1. If the variance of the data is greater, the residual mean deviance is likely to be greater than 1.

check_model(money.model.check)
```


```{r}
##foreign.diseases
foreign.diseases.model.check <- glm(code ~ 1 + Aspect + ClayPercentage, data = foreign.diseases.data, family = binomial())

check_model(foreign.diseases.model.check)
```


```{r}
##warfare
warfare.model.check <- lm(code ~ 1 + BulkDensity + OrganicCarbon + Slope + bio2 + bio15 + bio19, data = warfare.data)

check_model(warfare.model.check)
#Homoscedasticity looks good
#There seems to be a problem with the non-normality of residuals - this can be checked with a qqPlot (see below)
#Multicollinearity looks fine
#Must also check autocorrelation
#maybe one influential point

View(warfare.model.check)

#checking autocorrelation and mulitvariate normality (qqPlot)
check_autocorrelation(warfare.model.check)
qqPlot(warfare.model.check$residuals)
#autocorrelated results were detected (p = 0.012).
#there are outliers indicated by the qqPlot - these were investigated, then removed and the analysis re-run with the refined data (see below)

check_model(warfare.model.check)
```


```{r}
##warfare without outliers
extreme.warfare.values <- warfare.model.check$residuals
View(extreme.warfare.values)

#remove those societies with an outlier in their residual value
myVector <- warfare.model.check[['residuals']]
names.to.keep <-  names(myVector[myVector < 1.752 & myVector > -2.58])
new.warfare.data <- warfare.data[names.to.keep,]

##re-doing phylostep for new warfare data
phylostep (formula, starting.formula = NULL, data = new.warfare.data, phy = language.tree, method= "logistic_MPLE", direction = "both", trace = 0, btol = 100, log.alpha.bound = 4, start.beta=NULL, start.alpha=NULL, boot = 0, full.matrix = TRUE, k=2)

#new warfare model (incldues Aspect)
new.warfare.formula <- code ~ 1 + Aspect + BulkDensity + OrganicCarbon + Slope + bio2 + bio15 + bio19

new.LM.warfare <- phylolm(formula = new.warfare.formula, data = new.warfare.data, phy = language.tree, method = "logistic_MPLE", btol = 100, log.alpha.bound = 4)

#new warfare summary statistics
summary(new.LM.warfare)
View(new.LM.warfare$residuals)
View(new.LM.warfare$fitted.values)

##new warfare model check
new.warfare.model.check <- lm(code ~ 1 + Aspect + BulkDensity + OrganicCarbon + Slope + bio2 + bio15 + bio19, data = new.warfare.data)

check_heteroscedasticity(new.warfare.model.check)
check_autocorrelation(new.warfare.model.check)
check_outliers(new.warfare.model.check)
qqPlot(new.warfare.model.check$residuals)

check_model(new.warfare.model.check)
```


```{r}
##community.size
community.size.model.check <- lm(code ~ 1 + Aspect + BulkDensity + PETseasonality + OrganicCarbon + PhCaCL + bio15, data = community.size.data)

#model checks
check_heteroscedasticity(community.size.model.check) #Heteroscedasticity detected (p = 0.012)
check_autocorrelation(community.size.model.check) #independent residuals (p = 0.844)
check_outliers(community.size.model.check) #no outliers
check_collinearity(community.size.model.check) #OrganicCarbon = moderate (VIF = 5.67)
qqPlot(community.size.model.check$residuals)

check_model(community.size.model.check)
```


```{r}
##pop.density
pop.density.model.check <- lm(code ~ 1 + bio2 + bio15, data = pop.density.data)

#model checks
check_heteroscedasticity(pop.density.model.check) #Heteroscedasticity (p = 0.000)
check_autocorrelation(pop.density.model.check) #residuals independent (p = 0.146)
check_outliers(pop.density.model.check) #no outliers
qqPlot(pop.density.model.check$residuals)

check_model(pop.density.model.check)
```


```{r}
##pathogen.stress
pathogen.stress.model.check <- lm(code ~ 1 + PETDriestQuarter + PETseasonality + PhCaCL + Slope + bio2, data = pathogen.stress.data)

#model checks
check_heteroscedasticity(pathogen.stress.model.check) #homoscedastic (p = 0.149)
check_autocorrelation(pathogen.stress.model.check) #residuals independent (p = 0.062)
check_outliers(pathogen.stress.model.check) #no outliers
qqPlot(pathogen.stress.model.check$residuals)

check_model(pathogen.stress.model.check)
```


```{r}
##famine
famine.model.check <- glm(code ~ 1 + bio18, data = famine.data, family = binomial())

check_model(famine.model.check)
```


```{r}
##political.integration
political.integration.model.check <- glm(code ~ 1 + ClayPercentage + PETseasonality + PhCaCL + Slope + bio18, data = political.integration.data, family = binomial())

check_model(political.integration.model.check)
```


```{r}
##social.strat
social.strat.model.check <- glm(code ~ 1 + Aspect + PETseasonality + PhCaCL + bio15 + bio19, data = social.strat.data, family = binomial())

#model checks
check_model(social.strat.model.check)
#No problems with multicollinearity
#No problematically high influential variables (Cook's Distance)
check_outliers(social.strat.model.check) #No outliers

#give summary of the model output
summary(social.strat.model.check)
#Null deviance: 146.34 on 105 df
#Residual deviance: 133.37 on 100 df


check_model(social.strat.model.check)
```


```{r}
##violence
violence.model.check <- glm(code ~ 1 + Aspect + PETDriestQuarter + PETseasonality + OrganicCarbon + PhCaCL + bio15 + bio19, data = violence.data, family = binomial())

check_model(violence.model.check)
```


```{r}
##revenge
revenge.model.check <- glm(code ~ 1 + Aspect + BulkDensity + PETseasonality + PhCaCL + bio15, data = revenge.data, family = binomial())

#model checks
check_model(revenge.model.check)
#No problems with multicollinearity
#Some highly influential variables (Cook's Distance)!
check_outliers(revenge.model.check) #No outliers

#give summary of the model output
summary(revenge.model.check)
#Null deviance: 130.27 on 93 df
#Residual deviance: 115.45 on 88 df

check_model(revenge.model.check)
```


```{r}
##defence
defence.model.check <- glm(code ~ 1 + BulkDensity + PhCaCL +
bio15 + bio18 + bio19, data = defence.data, family = binomial())

#model checks
check_model(defence.model.check)
#Multicollinearity of bio18 = moderate!
check_collinearity(defence.model.check) #bio18 = 5.13 VIF
#No problematically high influential variables (Cook's Distance)
check_outliers(defence.model.check) #No outliers

#give summary of the model output
summary(defence.model.check)
#Null deviance: 125.02 on 92 df
#Residual deviance: 108.93 on 87 df
check_collinearity(defence.model.check)

check_model(defence.model.check)
```


```{r}
##disaster
disaster.model.check <- glm(code ~ 1 + Aspect + ClayPercentage + PETDriestQuarter + Slope, data = disasters.data, family = binomial())

check_model(disaster.model.check)
```









